\documentclass{agony}
\title{SPH4U: Dynamics Assignment}

\begin{document}
\thispagestyle{firstpage}
\textbf{\thetitle}

\begin{prob}
	Two heavy boxes, $m_{1}$ and $m_{2}$, lie stationary on different inclines, as shown.
	A rope runs over a pulley and connects the boxes.
	Mass 1, $m_{1}$ is 380kg. Assuming that each incline is frictionless and the system is in equilibrium, answer the
	following:

	\begin{enumerate}[(a)]
		\item \textbf{Draw FBD(s).}\\
		      \begin{minipage}[t]{.5\textwidth}
			      \underline{$m_{1}$ FBD:}
			      \input{./Diagrams/q1_fbd_a.tex}
		      \end{minipage}%
		      \begin{minipage}[t]{.5\textwidth}
			      \underline{$m_{2}$ FBD:}
			      \input{./Diagrams/q1_fbd_b.tex}
		      \end{minipage}

		\item \textbf{Find the magnitude of the tension in the cable.}\\
		      \begin{solution}
			      \vec{F_{net_{1x}}}            & = 0                                                              \\
			      0                             & = \vec{F_{g}} \sin 18\degree + \left\lvert F_{T}\right\rvert \\
			      0                             & = m_{1}g \sin 18 \degree + \left\lvert F_{T}\right\rvert         \\
			      0                             & = 380(-9.8) \sin 18 \degree + \left\lvert F_{T}\right\rvert      \\
			      \left\lvert F_{T}\right\rvert & = 1150.78\text{N}                                                \\
			      \left\lvert F_{T}\right\rvert & = 1200 \text{N}                                                  \\
		      \end{solution}
		      $\therefore$ The magnitude of tension in the string is 1200N.

		\item \textbf{Calculate the mass of \bm{$m_{2}$} needed to keep the system in equilibrium.}

		      \begin{solution}
			      F_{net_{2x}} & = 0                                         \\
			      0            & = \vec{F_{g}} \sin 33 \degree + 1150.78 \\
			      m_{2}        & = \frac{-1150.78}{-9.8\sin 33 \degree}      \\
			      m_{2}        & = 215.6\text{kg}                            \\
			      m_{2}        & = 220 \text{kg}
		      \end{solution}
		      $\therefore$ The mass of the second box ($m_{2}$) is 220kg.
	\end{enumerate}
\end{prob}

\begin{prob}
	A girl applies a 140N force to a 35.0kg bale of hay at an angle of 28\textdegree~above the horizontal.
	The force of friction acting on the bale is 55N.
	\begin{enumerate}[(a)]
		\item \textbf{What will be the horizontal acceleration of the bale?}
		      \begin{solution}
			      \vec{F_{net_{x}}} & = \vec{F_{a_{x}}} - \vec{F_{f}} & \vec{F_{net_{x}}} & = m\vec{a}                                 \\
			      & = 140 \cos 28\degree - 55       &\tab 66.81             & = 35\vec{a}                                \\
			      & = 66.81 \text{N[right]}         &\tab \vec{a}           & = 1.96\text{m/s\textsuperscript{2}[right]} \\
			      &                                 &\tab \vec{a}           & = 2.0\text{m/s\textsuperscript{2}[right]}
		      \end{solution}
		      $\therefore$ The horizontal acceleration is 2.0m/s\textsuperscript{2}[right]
		      \newpage
		\item \textbf{What is the coefficient of friction between the bale and the ground?}
		      \begin{solution}
			      \vec{F_{f}} &= \mu_{f}\vec{F_{N}}\\
			      55 &= \mu_{f}(\vec{F_{g}} - \vec{F_{a_{y}}})\\
			      \mu_{f} &= \frac{55}{35\times 9.8-140\sin 35\degree}\\
			      \mu_{f} &= 0.21
		      \end{solution}
		      $\therefore$ The coefficient of friction between the bale and ground is 0.21.
	\end{enumerate}

\end{prob}

\begin{prob}
	A 15kg box is pushed up a 35\textdegree~ramp.
	A force of 110N exists between the box and the ramp.
	\begin{enumerate}[(a)]
		\item \textbf{Draw an FBD showing a tilted coordinate system (label positive x-direction)}\\
		      \begin{center}
			      \input{./Diagrams/q3_fbd.tex}
		      \end{center}
		\item \textbf{What minimum force, \bm{$F$}, would be  necessary to move the box up the ramp at a constant speed?}
		      \begin{solution}
			      \vec{F_{net_{x}}} &= \vec{F_{a}} + \vec{F_{g_{x}}} + \vec{F_{f}}\\
			      0 &= \vec{F_{a}} + 15(-9.8)\sin 35\degree - 110\\
			      0 &= \vec{F_{a}} - 194.31\\
			      \vec{F_{a}} &= 194.31\text{N}
		      \end{solution}
		      $\therefore$ Since constant speed implies no acceleration ($\vec{F_{net}}=0$), the force required to move the box at constant speed must also result in 0 net force, meaning the minimum force required is 190N.
	\end{enumerate}
\end{prob}

\begin{prob}
	The apparatus shown in the diagram consists of a box of books connected by a string to a hanger plate loaded with masses.
	The mass, $m_{1}$, is for the box and books and the mass, $m_{2}$ is for the hanger with masses.
	The box is moving up the incline 35\textdegree~to the horizontal with constant speed.
	What is the coefficient of friction between the box and the incline?
\end{prob}
\vspace{-4mm}
\begin{minipage}[t]{0.3\textwidth}
	\textbf{\bm{$m_{1}$}:}
	\begin{flalign*}
		\vec{F_{net_{x}}} & = \vec{F_{g}}\sin35\degree + \vec{F_{T}}-\vec{F_{f}}   \\
		0                 & =0.45(-9.8)\sin35\degree+\vec{F_{T}}+\vec{F_{f}}       \\
		0                 & =-2.53+\vec{F_{T}}+\vec{F_{f}}                         \\
		\vec{F_{T}}       & =\vec{F_{f}}+2.53\text{N[uphill]}                    &
	\end{flalign*}
\end{minipage}%
\hspace{0.5cm}
\begin{minipage}[t]{0.3\textwidth}
	\textbf{\bm{$m_{2}$}:}
	\begin{flalign*}
		\vec{F_{net_{y}}} & = \vec{F_{g}}-\vec{F_{T}}         \\
		0                 & = 0.35 \times 9.8 - \vec{F_{T}}   \\
		\vec{F_{T}}       & = 3.43\text{N[up]}              &
	\end{flalign*}
\end{minipage}%
\hspace{0.2cm}
\begin{minipage}[t]{0.3\textwidth}
	\textbf{Set Equations for}\\
	\textbf{\bm{$F_{T}$} Equal:}
	\begin{flalign*}
		3.43        & = \vec{F_{f}}+2.53        \\
		\vec{F_{f}} & = 0.9\text{N[downhill]} &
	\end{flalign*}
\end{minipage}%
\vspace{4mm}
\begin{minipage}[t]{0.5\textwidth}
	\textbf{Solve for Friction Coefficient:}
	\begin{flalign*}
		\vec{F_{f}} & = \mu_{f}\vec{F_{N}}                  & \\
		0.9         & = \mu_{f}\times0.45(9.8)\cos35\degree   \\
		\mu_{f}     & = 0.25
	\end{flalign*}
\end{minipage}\\
\\
\\$\therefore$ The coefficient of friction between the box and incline is 0.25.
\end{document}